### 2020-02-18 readme.Rmd
lab notes

---
title: 'Read Me'
author: 'Florian Muthreich'
date: '2020/02/04'
output: pdf_document
---

#Lab-Notes
Samples 2020-02-13
Vial  Sample         Pinus  Betula  Salix  Spores
5*    A4-DL 138-140  +++    +              +++
4     D4-DL 161-163  ++     +              +++
4*    A5-UR  64- 66  ++     +       1      +++

1. add 1 cube to vial (~1cm^-3^)
2. wash with water -> dissolve sediment vortex
3. centrifuge 3min 3000rpm
4. decant; save decant
5. add 3 ml 1.43 SPT solution
6. vortex
7. centrifuge 3min 3000rpm
8. decant supernatant into new vial
9. repeat 5-7; decant into same vial
10. wash with water; save decant for checking
11. wash with water; save decant x2
12. wash residue with water; save in vial

2nd Run
Vial  Sample
2*    A4-DL 188-190
3*    A5-UR  72- 74
5     D4-DL 182-184
6     A4-DL 130-132
7*    A4-DL 163-165

3rd Run
Vial  Sample
4     A4-DL 175-177
5     D4-DL 149-151
6     A4-DL 145-147
7*    D4-DL 138-140