### 2021-07-06 manuscript.Rmd
boris comm to intro

---
title: 'Assessing variations in the chemistry of fossil- and modern-*Pinus* pollen'
date: '2020/02/04'
output:
  bookdown::pdf_document2:
    toc: false
    extra_dependencies: ["flafter"]
  bookdown::word_document2:
bibliography: references.bib
csl: biology-letters.csl
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(here)
source(here("R","packages.R"))
source(here("R","authors_and_affiliations.R"))
```

---
author: `r Authors`
---
`r Affiliations`

2500 words not including in tables title page references
2 figs and 2 tables
figures separate legends in text
separate cover letter

<!-- 
A  200
I  500
MM 400
R  500
D  800
-->

\newpage

# Abstract

The study of pollen-grain chemistry has gained increasing focus in palaeoecological research since it is possible that chemical variations observed in sporopollenin may be used for identification of pollen grains.
Fourier Transform Infrared (FTIR) spectroscopy has become one of the most popular methods for characterising the chemical variations present within pollen and spores.
Fossil pollen from three sediment cores and modern material was analysed by FTIR microspectroscopy using a new detector that allows simultaneous capture of multiple single grain spectra.
We also investigate the effect of two pollen extraction methods (acetolysis and density separation) on fossil pollen chemistry.
Our methods reproduce modern *Pinus* spectra well, compared to bulk analysis and other microspectroscopy approaches.
We report chemical differences between fossil and modern pollen, mainly lack of lipid and protein wavebands (1800 - 1500 cm^-1^), in addition to changes in the carbohydrate region (1200 - 1000 cm^-2^). 
Acetolysed fossil pollen was compared to non-acetolysed fossil pollen, which revealed differences in their chemical composition (e.g. changes in the 1200 - 1000 cm^-2^ region) that are most similar to reported acetolysed spectra from other taxa, suggesting that acetolysis alters fossil pollen chemistry.
We also report contaminations observed in non-acetolysed samples and the applications of high density separation extraction for FTIR microspectroscopy and palynology.

Keywords: FTIR, fossil pollen, Holocene, density separation, acetolysis, pollen chemistry

\newpage
<!-- READ DATA -->
```{r message=FALSE, warning=FALSE, include=FALSE}
source(here("R","ages.R"))
if (file.exists(here("data","output","DAL_spectra.rds"))) {
  DAL = readRDS(here("data","output","DAL_spectra.rds"))
  DAL$age = 0
} else {
  source(here("R","read_data.R"))
  read_data("DAL")
  DAL = readRDS(here("data","output","DAL_spectra.rds"))
}

if (file.exists(here("data","output","MFM_spectra.rds"))) {
  MFM = readRDS(here("data","output","MFM_spectra.rds"))
  MFM[MFM$depth=="A4UR_72-74", 'depth'] = "A5UR_72-74" #fixing typo in sample name
  MFM$age = fn_age(MFM$depth)
  MFM[MFM$treatment == "","treatment"] = "SPT"
} else {
  source(here("R","read_data.R"))
  read_data("MFM")
  MFM = readRDS(here("data","output","MFM_spectra.rds"))
}

if (file.exists(here("data","output","TSK_spectra.rds"))) {
  TSK = readRDS(here("data","output","TSK_spectra.rds"))
  TSK$age = fn_age(TSK$depth)
} else {
  source(here("R","read_data.R"))
  read_data("TSK")
  TSK = readRDS(here("data","output","TSK_spectra.rds"))
}

if (file.exists(here("data","output","fresh_spectra.rds"))) {
  FRE = readRDS(here("data","output","FRE_spectra.rds"))
} else {
  source(here("R","read_data.R"))
  read_data_fresh()
  FRE = readRDS(here("data","output","FRE_spectra.rds"))
}

FRE$Label = "FRE"
DAL$Label = "DAL"
TSK$Label = "TSK"
MFM$Label = "MFM"

outliers = c("DAL_rx1GZ", "DAL_2cxp7", "DAL_VzFi0", "DAL_ANdP0", "MFM_av67v", "MFM_OftM2", "MFM_xogJr", "MFM_hoG4y", "MFM_5UNIJ", "TSK_vvgOZ", "DAL_xgYjM", "TSK_m4hN6", "TSK_e04Go") #THese MFM ids dont exist anymore

DAL = DAL[!DAL$ids %in% outliers,]
TSK = TSK[!TSK$ids %in% outliers,]
MFM = MFM[!MFM$ids %in% outliers,]
```

<!-- PRE-PROCESS DATA -->
```{r message=FALSE, warning=FALSE, include=FALSE}
if (!hasName(DAL, "emsc")) {
  source(here("R","pre_process.R"))
}
```

<!-- PCA -->
```{r message=FALSE, warning=FALSE, include=FALSE}

```

\newpage

# Introduction
<!--Main Goal: Explore the chemical differences between fresh and untreated fossil pollen.-->

The study of plant microfossils (pollen and spores) is an integral part of palaeoecology and palaeobotany.
Pollen records are an important tool to reconstruct past floristic, vegetation and environmental changes [@magriNewScenarioQuaternary2006; @tzedakisVegetationChangeGlacialInterglacial1994; @huntleyMigrationSpeciesResponse1989; @noguesbravoCrackingCodeBiodiversity2018; @tinnerEcologyAbiesAlba2013] and provide insights into ecological processes which operate on decadal, centennial and millennial timescales (e.g. the post-glacial migration of tree taxa into new habitats following the Quaternary glacial maxima @gieseckeMovingFrontPopulation2005, @huntleyQuaternaryPalaeoecologyEcology1996, @tinnerCentralEuropeanVegetation2001).

The high preservation-potential of pollen grains in the geological record results from the presence of the resistant biopolymer sporopollenin, which is the main building block of the outer pollen-grain wall (exine).
Although the exact structure of sporopollenin remains unknown [ref], sporopollenin has been described as being similar to lignin, incorporating both aromatic compounds and cross-linked aliphatic components [@liMolecularStructurePlant2019].
Variations in the chemical composition of sporopollenin can reflect phylogenetic differences between plant groups [@nieropQuantitativeAnalysisDiverse2019]. 
There is also evidence that environmental signals are recorded in the chemistry of pollen grains, including temperature [@bagciogluMultiscaleVibrationalSpectroscopic2015], nutrients [@kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020],  and UV-B radiation [@rozemaUVBAbsorbanceUVB2001; @blokkerDevelopmentProxySurface2005; @fraserUVBAbsorbingPigments2011; @jardineSheddingLightSporopollenin2017].
Thus, in recent years, the study of pollen-grain chemistry has gained increasing focus in palaeoecological research since it is possible that chemical variations observed in sporopollenin may be used as proxies for past environmental conditions.
There has been a particular focus on using sporopollenin chemistry to reconstruct past changes in UV-B radiation [@seddonFossilPollenSpores2019] and for improving the taxonomic precision available for sediment records [@dellannaPollenDiscriminationClassification2009; @zimmermannAnalysisAllergenicPollen2016].

<!-- This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution  -->
<!-- Similarly, FTIR has become a popular method in the quantification of UV-B absorbing compounds to reconstruct changes in UV-B radiation from sediment sequences [ref]. -->
<!--MOVE TO DISCUSSION: Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods [ @mondolApplicationHighThroughputScreening2019; @zimmermannAnalysisAllergenicPollen2016; @guedesPollenRamanSpectra2014].
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen [@bagciogluMultiscaleVibrationalSpectroscopic2015; @kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020].
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes. -->

Successful application of pollen-chemistry techniques within palaeoecological research requires a thorough understanding of the factors influencing chemical variations in the modern environment and in the accurate estimation of the chemical spectra from pollen grains.
Any insights gained from studies on modern grains must also be transferable under fossil settings [@jardineImpactOxidationSpore2015; @jardineSporopolleninChemistryIts2021].
Fourier transform infrared (FTIR) spectroscopy has become one of the most popular methods for characterising the chemical variations present within pollen and spores in the modern environment.
Bulk FTIR techniques, which involve the analysis of large quantities of modern-pollen samples (approximately 1-2 mg) (e.g. @zimmermannCharacterizationPollenVibrational2010, @lahlaliATRFTIRSpectroscopyReveals2014), predominantly result with high-quality spectra with detailed biochemical ‘fingerprint’ of pollen, comprising specific signals of sporopollenins, lipids, proteins, carbohydrates and other chemical constituents [@bagciogluMultiscaleVibrationalSpectroscopic2015].
As a result, FTIR spectra obtained from bulk pollen samples have been used to identify chemical differences at the genus, and in some cases, species and population level [@muthreichChemicalVariationsQuercus2020; @zimmermannInfraredSpectroscopyPollen2014; @depciuchIdentificationBirchPollen2018; @diehnDiscriminationGrassPollen2020; @bagciogluMonitoringPlantEnvironment2017; @zimmermannHighthroughputFTIRSpectroscopy2017].
However, these approaches may be less relevant for pollen samples extracted from fossil sequences since the number of pollen grains analysed vastly exceeds the quantity which is practically available from a sediment-core sample.

An alternative approach involves the use of FTIR microscopes, where a FTIR spectrometer is coupled to a microscope. 
This enables acquisition of infrared spectra from a small area, with aperture covering several pollen grains, a single pollen grain, or even pollen grain substructure [@zimmermannVibrationalMicrospectroscopyEnables2015].
FTIR-microscopes are better suited to sediment samples as a result of the capacity to generate FTIR spectra from individual grains.
However, FTIR microspectroscopy is hindered by strong Mie scattering of IR radiation on pollen grains, resulting with anomalous features in their FTIR spectra [@zimmermannVibrationalMicrospectroscopyEnables2015; @blumelObservationMieRipples2018a; @lukacsRecoveryAbsorbanceSpectra2015].
Although the scattering issues can be tackled by a number of analytical and numerical solutions [@diehnDiscriminationGrassPollen2020; @zimmermannChemicalCharacterizationIdentification2018; @lukacsRecoveryAbsorbanceSpectra2015; @zimmermannAnalysisAllergenicPollen2016], FTIR measurement of individual pollen grains remains challenging.
Nevertheless, microscopic FTIR techniques have now been applied to both modern- and fossil-pollen grains preserved in sediments.
The approach has been successfully used to show that chemical differences are found between pollen at the genus and even species level [@julierChemotaxonomyToolInterpreting2016; @woutersenNovelApproachStudy2018;  @zimmermannChemicalCharacterizationIdentification2018; @jardineChemotaxonomyDomesticatedGrasses2019]. 
Moreover, microscopic-FTIR techniques have been used to observe chemical differences in pollen samples to reconstruct changes in UV-B absorbing compounds in different periods across Earth's history [@fraserUVBAbsorbingPigments2011; @jardinePollenSporesBiological2016; @jardineProxyReconstructionUltravioletB2019; @jardineSheddingLightSporopollenin2017].

One of key challenges faced by chemical palynologists using micro-FTIR and other approaches is the time taken to extract and analyse fossil-pollen samples.
Some FTIR microscopes are equipped with automated sample stages, enabling automated measurement of predetermined measurement positions.
Measurements are usually acquired by consecutive recording of spectra by a single element detector [@zimmermannAnalysisAllergenicPollen2016;@zimmermannChemicalCharacterizationIdentification2018; @zimmermannVibrationalMicrospectroscopyEnables2015].
When spectra are recorded with equidistant spatial steps (i.e. scanning measurement), the individual spectra can be assembled after the recording to obtain a chemical image [@jardineSporopolleninChemistryIts2021; @zimmermannVibrationalMicrospectroscopyEnables2015; @julierChemotaxonomyToolInterpreting2016].
However, recent developments in technologies related to FTIR allow for easier and faster capture of spectra of pollen grains. 
A focal plane array detector (FPA) is made of a 2D array of IR detectors, and allows true chemical imaging.
This enables simultaneous measurement of several single pollen grains from a relatively large surface area without the need to move the sample stage.
Therefore, it is of potentially high relevance for palaeoecological research.

When combined with imaging processing technology, it may allow for more rapid identification and extraction of FTIR spectra of pollen grains from sediment samples.
Whilst FTIR microspectroscopy with FPA detectors has been widely applied with microplastics [@primpkeAutomatedApproachMicroplastics2017] application on pollen samples was limited just to chemical characterization of isolated sporopollenin capsules  [@chiappePollenGrainsFunctionalized2017].
In particular, there has been no investigation into application of FPA detectors on measurement of fossil-pollen residues taken from sediment cores.

Palaeoecologists may also use a variety of chemical-processing techniques in order to speed-up analyses through the isolation and concentration of pollen from the sediment matrix [@ridingGuidePreparationProtocols2021].
One of the most common methods for sporopollenin extraction involves an acetolysis treatment, which, alongside a combination of treatments with hydrofluoric acid (HF) and hydrochloric acid (HCl), enables removal of organic materials, calcium carbonates and silicates, while preserving the morphological integrity of pollen grains [@erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989].
Acetolysis has also often been used in combination with FTIR on pollen as a way to standardise between modern and fossil samples, [@jardinePollenSporesBiological2016; @woutersenNovelApproachStudy2018] or to study the effects of acid extractions [@dominguezIsolationIntactPollen1998; @dominguezPollenSporopolleninDegradation1999; @loaderPreparationPollenStable2000; @schulzeosthoffPhenolsIntegratedCompounds1987; @jardineImpactOxidationSpore2015; @jardineSporopolleninChemistryIts2021].
Indeed, although acetolysis treatments at low temperatures did not create major chemical differences in *Lycopodium* spores [@jardineImpactOxidationSpore2015], it is uncommon for low-temperature treatments to be applied to fossil-sediment samples since the higher temperatures are required for oxidation of the non-pollen organic material.
Such treatments may then result in chemically altering effects on fossil-pollen composition [@jardineSporopolleninChemistryIts2021], whilst variations in sediment types mean that different combinations of sample-preparation techniques may be required for different sediment types [@ridingGuidePreparationProtocols2021].
Given the large variations in FTIR spectra observed across taxa [@diehnCombiningChemicalInformation2020; @muthreichChemicalVariationsQuercus2020], and the variety of sediment types even from Holocene sequences in Europe, more studies are needed to investigate the potential effects of the chemical preparation procedures used to isolate pollen grains from different pollen types.

In this study we aimed to explore the use of a FTIR imaging to obtain IR spectra from fossil-pollen *Pinus* grains.
We develop a workflow which enables comparison of fossil-pollen grains taken from three sediment-core locations (Table \@ref(tab:sites)), and compare the IR spectra from both modern- and fossil-pollen samples.
Pollen was extracted from the sediment cores using two different approaches, namely density separation using sodium polytungstate and acetolysis, followed by FTIR microscopy imaging.
Thus, in addition to testing the feasibility of the FTIR imaging approach on different sediment types, we were were also able to investigate how the IR spectra of the *Pinus* pollen vary as a result of two different extraction procedures.

<!-- Chemical treatment... -->
<!-- SPT density separation as an alternative to acetolysis by investigating the chemical alteration of both extraction methods to three fossil cores. -->
<!-- With this, we aim to investigate changes in fossil pollen chemistry, due to: i. differences between modern and fossil material, ii. extraction method and iii. effect of diagenesis (age of fossil pollen). -->
<!-- For this purpose we use an FTIR microscope with FPA detector to capture spectra of modern and fossil *Pinus sylvestris* pollen. -->
<!-- <!-- -->
<!-- The effect of treatment with SPT on pollen chemistry is tested on modern material to establish if any chemical changes occur. -->
<!-- -->

<!-- RQs -->
<!-- i. Differences between cores -->
<!-- ii.i Difference between acetolysed and fossil -->
<!-- ii.ii Difference between fresh and fossil grains. -->
<!-- iii effect of depth/age on chemistry -->

<!-- iv effect of orientation on chemistry -->

<!-- #Alistair rewritten paragraphs- maybe move to methods -->

<!-- An alternative to chemical extraction is density separation using e.g. sodium polytungstate (SPT), which is used to isolate pollen from minerogenic sediments [REFS] or as preparations for pollen isotope studies [REFS]. -->
<!-- Density separation may be a an alternative to acid extractions for pollen studies concerned with chemical alteration of sporomorph chemistry. -->
<!-- SPT is used to adjust the density of sample suspension and facilitace separation of less dense particles (e.g. pollen and other microfossils) over mineral particles. -->
<!-- SPT is removed from the sample with several washes after extraction and any contamination of the sample is easily identifiable by characteristic wavebands in the spectra.  -->

\newpage

# Methods and Materials

```{r sites, echo=F, results='asis'}
sites = data.frame(name = c("Tiefer See",
                            "Dalmutladdo",
                            "Meerfelder Maar"),
                   abbreviation = c("TSK",
                                    "DAL",
                                    "MFM"),
                   age_range = c("1932 - 2015 AD",
                                 "9836 - 595 BP",
                                 "13120 - 11400 BP"),
                   description = c("varved sediment, formed during the last last glacial period, Mecklenburg-Vorpommern, Germany",
                                   "formed after the last deglaciation, near Tromsø, Norway ",
                                   "varved sediment, formed in a volcanic crater in the Westeifel, Germany"),
                   publication = c("(ref:tsk)",
                                   "(ref:dal)",
                                   "(ref:mfm)"))

knitr::kable(sites, booktabs = T, format = "latex", 
             caption = "",
             col.names = c("Name", "Abbr.", "Age", "Description", "Publ.")) %>%
  column_spec(4, width = "5cm")
```

(ref:tsk) @theuerkaufEffectsChangesLand2015
(ref:dal) @bjuneHoloceneVegetationClimate2004
(ref:mfm) @brauerHighResolutionSediment1999

The pollen was extracted from three sediment cores: Dalmuttlado in Norway [@bjuneHoloceneVegetationClimate2004], Tiefer See [@theuerkaufEffectsChangesLand2015]  and Meerfelder Maar [@brauerHighResolutionSediment1999] in Germany.
The three sediment cores were selected because they contain *Pinus sylvestris*-pollen types, spanning three different periods in the late Quaternary (Table \@ref(tab:sites)).
*Pinus* pollen in these cores has been described as most likely representing *Pinus sylvestris* in Dalmutladdo and Tiefer See, while it is described as *Pinus* in Meerfelder Maar record.
In this study we focus on *Pinus* pollen for two reasons.
First, *Pinus* spp. is a shade-intolerant, high pollen producer, and so a number of studies have proposed that *Pinus* spp. pollen may be a useful taxon to use as a biological indicator for reconstructing past changes of UV-B radiation [@jardineProxyReconstructionUltravioletB2020; @willisQuantificationUVFlux2011; @seddonPollenchemistryVariationsElevation2021].
Second, although prominent in Quaternary sediment records from Europe, and despite the highest number of species amoing conifers [110 species, @gernandtPhylogenyClassificationPinus2005], *Pinus* is currently mostly only split into two different pollen types using fossil pollen, reflecting the two phylogenetic lineages [@beugLeitfadenPollenbestimmungFur2004; @gieseckePostglacialChangeFloristic2019].
Thus, identifying chemical differences in fossil *Pinus* pollen may also allow for identification of crytpic *Pinus* species from the pollen types observed in the fossil record.

Prior to chemical characterisation using FTIR, all sediment samples were processed in two ways.
First, the samples were extracted using density separation using dissolved sodium polytungstate (SPT) [@kampsAgentSeparationDissolved1985] at 1.43 g cm^-3^ as a separation medium.
During density separation a known volume of SPT (density 1.43 g cm^-3^) solution was added to the sample and homogenized/vortexed, followed by centrifugation.
After centrifugation, particulates heavier than 1.43 g cm^-3^ accumulate of the bottom of the sample vial, whilst the lighter fraction was decanted and collected.
Note, that there is no agreement on the density of the pollen exine, with reported values varying between <1.4 g cm^-3^ [@faegriTextbookPollenAnalysis1989; @flenley1971measurements; @regnellPreparativeCentrifugationNew1996], 1.88 cm^-3^ [@nakagawaDensemediaSeparationMore1998] and 2.2 cm^-3^  [@juvigneDensiteExinesQuelques1973]. 
In our case the lighter fraction (<1.4 g cm^-3^) contained majority of the pollen grains, after checking both fractions.
Sample residues were then washed three times in using distilled H~2~O to remove any excess SPT, resulting in sample residues with a higher pollen concentration compared to other organic and inorganic materials on the slide.
Since SPT is a chemically inert salt, it was expected that, after washing, this treatment should have minimal chemical effect on the chemical spectra inferred from the fossil-pollen grains.

We then applied an acetolysis treatment [e.g. @erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989] to one subset of the samples following washing and extracting using SPT.
The acetolysis was performed at 105&deg;C for 5 minutes.
Incorporating the acetolysis step on this analysis allows us to compare the effects of pollen-grain spectra from both treated and untreated pollen grains.

The modern *Pinus sylvestris* pollen samples were collected from locations close to Bergen, (Norway) in spring 2019 and Innsbruck, (Austria) in spring 2020.
Pollen was sampled from four to six pollen cones and air-dried at room temperature for 24h.
For this study, pollen from two trees at each location were selected and analysed.

## FTIR measurements
The spectra were measured with a Vertex 70 FTIR spectrometer connected to a Hyperion 3000 IR microscope (Bruker Optik, Ettlingen, Germany) equipped with a focal-plane-array (FPA) nitrogen-cooled mercury-cadmium-telluride (MCT) detector and a 15x magnification objective.
The FPA-MCT detector allows the capture of FTIR images, where each pixel is a full spectra, instead of point measurements.
For this study images of 128x128 pixels were taken and each spectra was recorded with 64 scans.
The accompanied visible images were obtained by a charge coupled device (CCD) camera coupled to the microscope.
Single-grain FTIR spectra were extracted from the FTIR images, by averaging all pixel spectra in the image area of the pollen grain. 
Data acquisition and instrument control was performed using the OPUS 7.5 software (Bruker Optik GmbH, Germany).

The extracted pollen was applied to ZnSe slides and a background spectra was measured before each sample on an empty area of the slide.
The pollen treated with SPT and acetolysis were suspended in water, applied to the slide and then dried out before measurement under a fume hood.
The extracted sediment samples contained other small material than pollen (e.g. diatom fragments, organic material fragments, etc).
We recorded several "blank" spectra of dried residue that formed over the dried pollen residue (see also SM).

## Spectra preprocessing and analysis

The spectra were preprocessed before analysis consisted of smoothing and calculation of second derivative and normalization.
We used the Savitzky Golay (SG) algorithm [@savitzkySmoothingDifferentiationData1964] as implemented by the extended multiplicative signal correction (EMSC) package in R [@lilandPackageEMSC2017] and used second degree polynomial and window size five as parameters.
The second derivative spectra was then normalized using a EMSC model with a linear component [@lilandPackageEMSC2017].
Preprocessing with SG and EMSC has been shown to improve the performance of multivariate analyses, such as principal component analysis (PCA) or models in the partial least squares regression (PLSR) family [@woutersenNovelApproachStudy2018; @zimmermannOptimizingSavitzkyGolayParameters2013].

We used two PLSR models to explore the chemical differences between modern and fossil grains (model 1), and acetolysed and non-acetolysed fossil grains (model 2).
For these models the explanatory variables were the pre-processed spectra and the response variables were hot-one encoded response variables (Model 1: modern/fossil; Model 2: acet./non acet.).
By using PLSR we aimed to maximize the distance between responses and identify chemical differences between the responses by looking at the regression coefficients.

\newpage

# Results
<!-- Our aim with this study was to explore the use of a new type of detector (FPA-MCT) in combination with a FTIR microscope to explore the chemical variations of fossil pollen. -->
<!-- Our goal was also to explore the effects of two different sample-processing methods on fossil pollen and compare the spectra we obtained with modern samples. -->

```{r meanspectra, echo=FALSE, fig.cap="Mean spectra summarised by core and chemical treatment", message=FALSE, warning=FALSE, out.width="90%"}
knitr::include_graphics(here("figures", "mean_plots.png"))
```

\newpage

```{r ordinations, echo=FALSE, fig.cap="a. plsr ordination scores all samples diff fresh vs fossil b. regression coef; component 1 c. regression coef component 2; d plsr ordination scores only samples from cores diff spt vs acet; e. regression coef component 1; f. regression coef component 2", message=FALSE, warning=FALSE, out.width="100%"}
knitr::include_graphics(here("figures", "ord_plots.png"))
```

\newpage
## Chemical differences between pollen samples
Our analyses revealed clear differences between modern- and fossil-pollen grains (Fig. \@ref(fig:meanspectra)).
Although the prominent and characteristic sporopollenin peaks (1600, 1510 and 1170 cm^-1^) are visible in both the modern and fossil pollen, other parts of the spectra show substantial differences.
Most notably, the lipid peak at 1745 cm^-1^, caused by triglycerides, is not visible in the fossil-pollen grains, neither in the non-acetolysed and the acetolysed fossil pollen.
There are also several differences in the "fingerprint" region (1800 - 1500 cm^-1^) and carbohydrate region (1250 - 1000 cm^-1^) between modern- and fossil-pollen grains.

These chemical differences are also clearly demonstrated in the results of the PLSR (Fig \@ref(fig:ordinations) a).
Fossil pollen are clearly separated from the modern pollen along the first component of the PLSR model.
The regression coefficients reflect our observations from the visual inspection and show that lipid and protein peaks (1745 and 1800 - 1500 cm^-1^) are no longer present.
There are also substantial changes in the carbohydrate region (1250 - 1000 cm^-1^) in the fossil pollen compared to the modern (Fig \@ref(fig:ordinations) b and c).
The region 1250 - 1000 cm^-1^ is now dominated in acetolysed fossil pollen by two peaks and its shoulders at 1170 and 1040 cm^-1^.

In addition to the differences between fresh and fossil pollen samples, our analysis revealed large differences in chemical variations as a result of the acetolysis treatment on the fossil *Pinus* spp. pollen grains.
For example, acetolysed pollen showed two dominant peaks in the carbohydrate region at 1190 - 1170 and 1040 cm^-1^, which are not visible in the modern spectra or the non acetolysed pollen, and when using the second PLSR model, the acetolysed pollen is clearly separated from the non-acetolysed pollen, with a much greater distance in the ordination space between samples (Fig. \@ref(fig:ordinations) d).
This change in the 1190 - 1170 cm^-1^ region is unique to acetolysized pollen and had higher regression coefficients that the surrounding wavenumbers (Fig. \@ref(fig:ordinations)).
The 1040 cm^-1^ peak also had high importance for the PLSR model differentiating acetolysed and non-acetolysed fossil pollen (Fig. \@ref(fig:ordinations) e and f).

## Contamination effects

```{r SM, echo=FALSE, fig.cap="Image captured from FTIR microscope of sample slide Dalmutladdo 60 cm depth. Pinus pollen is visible as well as debris and dried residue, presumably BSi.", message=FALSE, warning=FALSE, out.width="90%"}
knitr::include_graphics(here("figures", "SM_BSI.png"))
```

In addition to the clear differences between samples, we also observed two potential contamination effects from the sediment core processing.
First, in the Tiefer See non-acetolysed fossil pollen samples (Fig. \@ref(fig:meanspectra)) there was a broad peak at 1630 cm^-1^. 
This 1630 cm^-1^ peak was responsible for part of the separation between Tiefer See non-acetolysed samples from the other cores, but does not change the larger differentiation of acetolysed Tiefer See samples (without the 1630 cm^-1^ peak) from the other cores, and the peak very likely indicates the presence of SPT [see spectra from @SodiumMetatungstateHydrate].
Thus, it is possible that the three washes with water after decanting of the pollen fraction were not enough in this case, but absence of this peak in the other spectra seem to suggest that further washes may remove successfully remove SPT fully.

Second, we also observed a broad peak at 1100 - 1090 cm^-1^ in the non-acetolysed spectra of all three sediment cores (Fig. \@ref(fig:meanspectra)).
Comparisons of FTIR spectra from biogenic-silica spectra from diatoms [@alipourInfraredMicrospectroscopyBionanomaterials2016] and hydrated silica spectra [@SilicaATRIRSpectrum] reveal that the peak at 1090 cm^-1^ in the non acetolysed samples is most likely biogenic silica, which may have precipitated onto the onto ZnSe slides as the sediment residues being prepared for analysis. 
During the measurements with the microscope, we applied our extracted pollen as droplets onto ZnSe slides and waited for the water to evaporate and we observed a crystalline layer over the same area that covered all particles and pollen (Fig. SM \@ref(fig:SM)).
Interestingly, it appeared that the acetolysis treatment removed this peak, which may be because of the extended washing with several chemicals and water during standard acetolysis.

# Discussion

## Extraction of *Pinus* FTIR spectra using the FPA detector
In this study we investigated the use of an FPA detector in combination with a FTIR microscope that allows capture of several spectra in order to characterise the chemistry of pollen grains from a sediment sample.
The approach has the potential to be an improvement on previous single-grain measurements using FTIR, where each grain is required to be targeted individually, or even manually extracted from the sediment matrix using a micro-pipette.
The FPA-MCT detector on our FTIR microscope allows simultaneous capture of a larger area of the sample (XXX x XXX µm ), compared to other detectors (e.g. 100 x 100 µm @jardineChemotaxonomyDomesticatedGrasses2019, 50 x 50 ?µm? @zimmermannChemicalCharacterizationIdentification2018).
An image with a spectra for each pixel is created, from which spectra of single grains were extracted.

In general the FTIR spectra obtained using this approach created similar spectra to spectra of *Pinus* pollen from bulk approaches, and the method was also sensitive to differences in chemical treatments on the fossil-pollen grains.
For example, the spectra obtained from our modern samples are very similar to *Pinus sylvestris* spectra measured using FTIR-bulk sample methodologies [@bagciogluMultiscaleVibrationalSpectroscopic2015; @zimmermannCharacterizationPollenVibrational2010] and FTIR microscopy using single-point measurements [@zimmermannChemicalCharacterizationIdentification2018].
The same characteristic peaks described above are visible in all modern spectra captured with bulk and FTIR microscopy, e.g. 1745, 1515, 1170 and 1250 - 1000 cm^-1^, and these results build on previous studies which suggest that overall potential for the use of the FPA detector in pollen research **[add references]**.
Thus, moving forward we see potential in the application of the FPA detector in chemical palaeoecology.

## Variable effects of acetolysis
Another important result in our study is the large differences in FTIR spectra between acetolysed and non-acetolysed grains.
The presence of the peak in the 1170 cm^-1^ shoulder in the spectra from the acetolysed pollen grains suggests an overall change in the pollen chemistry of the sample.
The presence of these two dominant peaks may be related to the removal and or alteration of grain-wall carbohydrates (cellulose etc), as well as changes to chemical composition of sporopollenin due to the acetolysis treatment.
Since the 1190 cm^-1^ peak is not visible in any of the samples from the non-acetolysed pollen, this likely indicates that it is processes related to the acetolysis treatment that are creating this new peaks in the FTIR spectra.

Interestingly, when comparing our acetolysed *Pinus* spectra to other acetolysed fossil *Pinus* pollen, we observed some differences.
For example, @jardineProxyReconstructionUltravioletB2020 extracted *Pinus* pollen from a varved sediment core using acetolysis and HF to study the UV-B response in *Pinus* pollen chemistry.
But while their spectra are similar to ours in the 1800 - 1500 cm^-1^ region, with similar prominent peaks at 1515 cm^-1^, 1700 and 1600 cm^-1^, there are major differences in the region between 1500 - 1300 cm^-1^.
The latter region is fairly flat in our spectra, while a strong peak is visible at 1450 in @jardineProxyReconstructionUltravioletB2020 spectra.
Lastly, the region between 1300 - 900 cm^-1^ in our spectra mainly consists of two large peaks at 1200 and 1040 cm^-1^, while the *Pinus* pollen from @jardineProxyReconstructionUltravioletB2020 shows large peaks at 1250, 1150, 1100 and 1010.

Conversely, we found greater similarities between the FTIR spectra from the acetolysed *Pinus* pollen from our study to  acetolysed pollen grains of other pollen types.
For example, the carbohydrate region at 1200 - 1000 cm^-1^ in our acetolysed *Pinus* pollen samples is most similar to an acetolysed *extant* Poaceae set of @jardineSporopolleninChemistryIts2021, and in addition to acetolyzed herbarium *Nitrariaceae* pollen [@woutersenNovelApproachStudy2018], where two peaks at 1040 and 1190/1170 are visible and similar to our samples.
However, spectra from a fossil Poaceae data set (sampled from the the Amazon Fan and Western Amazon; AF: Na~4~P~2~O~7~, density sep., lignite, WA: HCl, Na~4~P~2~O~7~, density sep., acetolysis) [@jardineSporopolleninChemistryIts2021] is quite different in composition to both our fossil pollen, and any extant acetolysed-pollen samples, since the large peak at 1600 and 1190 peak are not clearly visible.

Taken together, the differences in the effects of acetolysis across different taxa and sample types suggest that the chemical extraction methods have a large effect on the chemistry of pollen, both modern and fossil origin.
Direct comparisons between spectra from different spectrometers are difficult, because peak heights are not always directly comparable, but these results indicate that further research is required to understand the reasons between these differences between the acetolysed *Pinus* pollen from across the sample sets.
This is important to consider going forward for the application of IR methods in palaeoecology, where extraction methods are often dependent on sediment type.
For example, it would be useful to explore the effects of these extraction methods on modern and fossil material side-by-side, to investigate if there are methods that can bridge the transition from modern to fossil pollen.
Chemical treatments that can mimic diagenesis will be crucial to understanding the chemical changes of pollen under diagenesis and help to create species specific reference collections of artificial fossilized modern pollen grains.
This would enable us to compare fossil pollen spectra with fossilized pollen spectra of known species for identification.

<!-- @jardineImpactOxidationSpore2015 showed limited effects on pollen chemistry of cold acetolysis if it was shorter than 10 minutes. -->
<!-- Hot acetolysis (90&deg;C), which is standard procedure in palynology [@erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989], had an effect from the first minute.  -->

The chemical effects of acetolysis that identified across studies suggests that it may be preferable to use an alternative approach to extract fossil pollen grains from sediment samples.
Given the limited chemical influence on the fossil pollen grains, we believe that density separation is a viable alternative to acetolysis and other acid or oxidizing extractions [@caffreyUseLithiumHeteropolytungstate2013; @leipeTestingPerformanceSodium2019; @regnellPreparativeCentrifugationNew1996], although given the fact that an SPT peak was observed in one of the cores, the results in this paper clearly demonstrate the need to identify efficient dilution processes that may need to be tailored to specific sediment types.
Similarly, since the reported density of pollen is variable and may be dependent on age of the sample [@faegriTextbookPollenAnalysis1989; @flenley1971measurements; @regnellPreparativeCentrifugationNew1996; @nakagawaDensemediaSeparationMore1998; @juvigneDensiteExinesQuelques1973], it is possible that specific extraction densities and procedures may be required from core to core.

Since our non-acetolysed samples are characterised by a BSi peak in the FTIR spectra, the density-separation approach also appears to introduce an additional important effect.
Although the BSi peak does not obscure any of the sporopollenin peaks in the FTIR spectra, it does obscure the region 1110 - 1070 cm^-1^.
The quantity of biogenic silica may vary between different sediment cores from different lakes and regions, which may influence the severity of the interference caused by BSi. 
For pollen-chemistry research this may present a challenge if we are aiming to directly compare spectra from unprocessed pollen samples since it may create differences in the spectra beween modern and fossil grains.
A solution would be to apply correction procedures during pre-processing, or explore extended-washing iterations after extraction of the pollen to dilute BSi.

## Future developments
In this study we focused on the chemistry of *Pinus* pollen, because of it's importance in studies related the UV-B-related research and due to its high ecological importance in Quaternary sequences, particularly in the northern hemisphere.
*Pinus* pollen is also relatively large, which prevents any scattering problems found with smaller grains and made for easy identification of pollen granis in from the images created in our experimental set up.
However, the quality of the optical images captured by our microscope was much worse than convential LM, which made it difficult to identify smaller pollen grains on the slide from other organic debris, e.g. *Betula*/*Corylus*, *Alnus* or *Quercus*.
An objective with higher magnification (20x) may improve this aspect of using a FTIR microscope.
Since FTIR measurements of small particles (5-25 µm) introduces scattering artifacts into the spectra [@lukacsRecoveryAbsorbanceSpectra2015; @zimmermannVibrationalMicrospectroscopyEnables2015], sample processing and computational methods exist to minimize or avoid scattering artifacts for these smaller taxa using FTIR.

Finally, a potential future area of research would then be to investigate the possibility for automation to streamline the spectral-collection process. 
Although in this study the spectra were extracted by hand, approaches using deep learning have shown to be capable of extracting objects from larger images, using deep neural networks specialised on image segmentation/object location [@renFasterRCNNRealTime2015].
Conceptually the problem of locating pollen on a FTIR image is similar to locating an object on a optical image.
@primpkeAutomatedApproachMicroplastics2017 devloped an automated approach to identify microplastic with a FPA detector.
@kangSinglecellClassificationFoodborne2020 used deep neural network to classify bacteria based on images and spectral information, while @grosserueschkampSpatialMolecularResolution2017 used random forests for segmentation and identification of malignant tumor in histological thin sections using FTIR microscope equipped with a FPA detector (715 x 715 µm).
These applications from other disciplines with a similar task show the potential of FPA detectors for chemical palynology using FTIR microscopy.

<!-- After testing multiple densities, 1.44 g cm^-3^ yielded the most pollen from our samples. -->
<!-- This density may be optimal for *Pinus* in our samples, but other pollen may have been discarded with the heavier fraction. -->
<!-- Our results show that there is a need to refine extraction methods, which minimize chemical alteration to the pollen grain. -->
<!-- A refinement of the right extraction medium density in combination with sieving of the samples to remove non pollen organic material may be one approach. -->
<!-- @caffreyUseLithiumHeteropolytungstate2013 and @leipeTestingPerformanceSodium2019 have reported similar counts of pollen grains to conventional extraction methods using density separation, but also mention that ideal extraction density varies from core to core.  -->

## Summary
Fossil pollen from three sediment cores and modern material was analysed by FTIR microspectroscopy using a new detector that allows simultaneous capture of multiple single grain spectra.
Our methods reproduces modern *Pinus* spectra well, compared to bulk analysis and other microspectroscopy approaches.
Acetolysed fossil pollen was compared to non-acetolysed fossil pollen, which revealed differences in their chemical composition (e.g. changes in the 1200 - 1000 cm^-2^ region) that are most similar to reported acetolysed spectra from other taxa, suggesting that acetolysis alters fossil pollen chemistry.
We also report contaminations observed in non-acetolysed samples and the applications of high density liquid extraction for FTIR microspectroscopy and palynology.

## I haven't found a way to integrate these ideas in there. Perhaps we don't need them? I think perhaps we could leave it out
In addition, there is some differentiation between the cores.
Samples from Tiefer See are more distinct from samples of the other two cores, which show more similar scores in the ordination.
One interpretation of this is that *Pinus* pollen from DAL and MFM are chemically quite similar, while TSK *Pinus* pollen is different in chemical structure.
Since it is likely that the main contributions to the pollen some in the TSK core is *Pinus sylvestris*, it is likely that the same species is represented by all three cores, and one of the main differences between the TSK pollen and the other two locations is sample age.
Whilst the TSK samples are from the past 80 years, MFM covers the YD/Holocene transition and DAL samples are from throughout the Holocene, with the youngest sample likely more than 100 years older than TSK pollen.

This finding could suggest a chemical difference in fossil-pollen grains that is persistent even after acetolysis, which may be explained by the age of the pollen.
Interestingly, @jardineSporopolleninChemistryIts2021 also found differences in FTIR spectra with different sample ages, although they found that that modern-pollen spectra had more in common with spectra from Quaternary pollen, compared to pollen from deeper time samples.
Further studies on pollen from different sediment types at fixed age/depths may provide a clearer answer on the effects of sediment type and age on pollen chemistry.

# References